---
title: Using Priming to Uncover the Organization of Syntactic Representations in Neural Language Models
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(RColorBrewer)
library(ggpubr)
library(reshape2)
library(tidyverse)

base_size <- 17

theme_set(theme_bw(base_size = base_size) +
            theme(panel.grid.major=element_blank(),
                  panel.grid.minor=element_blank(),
                  legend.key.size=unit(1.5, 'lines'),
                  legend.spacing=unit(0, 'in'),
                  axis.text.x=element_text(size=rel(1.1)),
                  axis.title.x = element_text(size = base_size)))

```





**Defining functions**

```{r}

data_summary <- function(data, varname, groupnames, ci = TRUE){
   length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

  data_sum <- data %>% group_by(.dots = groupnames) %>% summarise(N = length2(get(varname)), mean = mean(get(varname), na.rm = TRUE), sd = sd(get(varname), na.rm = TRUE)) %>% rename(!!varname := mean)
  

  if(ci==FALSE){
    return(data_sum)
  }
  else{
    data_sum$se <- data_sum$sd / sqrt(data_sum$N)

  ciMult <- qt(0.95/2 + .5, data_sum$N-1)
  data_sum$ci <- data_sum$se * ciMult
 return(data_sum)
  }
}

```


```{r}

#for dlinking analysis separate

# setwd("/Users/Debasmita/Desktop/RNN-Priming")
# post_summary <- readRDS('link_pre_list_summary_no.unk.Rda')
# shuffled_post_summary <- readRDS('shuffled_link_pre_list_summary_no.unk.Rda')
# linked <- c('linkedwh')
# post_summary$Dlinked <-ifelse(post_summary$adapt %in% linked, 'yes', 'no')
# shuffled_post_summary$Dlinked <-ifelse(shuffled_post_summary$adapt %in% linked, 'yes', 'no')
# 
# agg <- aggregate(post_summary$post_surp,by=list(post_summary$Dlinked), FUN=mean)
# shuffled_agg <- aggregate(shuffled_post_summary$post_surp,by=list(shuffled_post_summary$Dlinked), FUN=mean)
# 
# post_summary$mean <- ifelse(post_summary$Dlinked=='yes', 7.354098, 7.474417)
# shuffled_post_summary$mean <- ifelse(shuffled_post_summary$Dlinked=='yes', 8.905919, 7.866279)
# 
# write.csv(post_summary, "ordered_dlink.csv", row.names=FALSE)
# write.csv(shuffled_post_summary, "shuffled_dlink.csv", row.names=FALSE)

```

## Set up
### Load in data from trained models
```{r, cache = TRUE}

setwd("/Users/Debasmita/Desktop/RNN-Priming/")
pre_summary <- readRDS('pre_list_summary_no.unk.Rda') 
post_summary <- readRDS('post_list_summary_no.unk.Rda')

shuffled_pre_summary <- readRDS('shuffled_pre_list_summary_no.unk.Rda') 
#shuffled_pre_summary <- shuffled_pre_summary[(shuffled_pre_summary$sentid <30)&(shuffled_pre_summary$sentid !=1)&(shuffled_pre_summary$sentid !=11)&(shuffled_pre_summary$sentid !=24)&(shuffled_pre_summary$sentid !=22)&(shuffled_pre_summary$sentid !=20),]
#del <- ("sentid")
#shuffled_pre_summary <- shuffled_pre_summary[,!(names(shuffled_pre_summary) %in% del)]
#shuffled_pre_summary$sentid <-0
#count <- 1
# for (row in 1:nrow(all20)){
#   if(count<1001){shuffled_pre_summary[row, "sentid"] = 24}
#   if(count<961){shuffled_pre_summary[row, "sentid"] = 23}
#   if(count<921){shuffled_pre_summary[row, "sentid"] = 22}
#   if(count<881){shuffled_pre_summary[row, "sentid"] = 21}
#   if(count<841){shuffled_pre_summary[row, "sentid"] = 20}
#   if(count<801){shuffled_pre_summary[row, "sentid"] = 19}
#   if(count<761){shuffled_pre_summary[row, "sentid"] = 18}
#   if(count<721){shuffled_pre_summary[row, "sentid"] = 17}
#   if(count<681){shuffled_pre_summary[row, "sentid"] = 16}
#   if(count<641){shuffled_pre_summary[row, "sentid"] = 15}
#   if(count<601){shuffled_pre_summary[row, "sentid"] = 14}
#   if(count<561){shuffled_pre_summary[row, "sentid"] = 13}
#   if(count<521){shuffled_pre_summary[row, "sentid"] = 12}
#   if(count<481){shuffled_pre_summary[row, "sentid"] = 11}
#   if(count<441){shuffled_pre_summary[row, "sentid"] = 10}
#   if(count<401){shuffled_pre_summary[row, "sentid"] = 9}
#   if(count<361){shuffled_pre_summary[row, "sentid"] = 8}
#   if(count<321){shuffled_pre_summary[row, "sentid"] = 7}
#   if(count<281){shuffled_pre_summary[row, "sentid"] = 6}
#   if(count<241){shuffled_pre_summary[row, "sentid"] =5}
#   if(count<201){shuffled_pre_summary[row, "sentid"] = 4}
#   if(count<161){shuffled_pre_summary[row, "sentid"] = 3}
#   if(count<121){shuffled_pre_summary[row, "sentid"] = 2}
#   if(count<81){shuffled_pre_summary[row, "sentid"] = 1}
#   if(count<41){shuffled_pre_summary[row, "sentid"] = 0}
#   count <- count + 1
# }
shuffled_post_summary <- readRDS('shuffled_post_list_summary_no.unk.Rda')
#shuffled_post_summary <- shuffled_post_summary[(shuffled_post_summary$sentid <25),]

drop <- c("N", "sd", "se", "ci")

pre_summary <- pre_summary[,!(names(pre_summary) %in% drop)]
colnames(pre_summary)[colnames(pre_summary)=="pre_test"] <- "test"

post_summary <- post_summary[,!(names(post_summary) %in% drop)]
colnames(post_summary)[colnames(post_summary)=="post_test"] <- "test"

shuffled_pre_summary <- shuffled_pre_summary[,!(names(shuffled_pre_summary) %in% drop)]
colnames(shuffled_pre_summary)[colnames(shuffled_pre_summary)=="pre_test"] <- "test"

shuffled_post_summary <- shuffled_post_summary[,!(names(shuffled_post_summary) %in% drop)]
colnames(shuffled_post_summary)[colnames(shuffled_post_summary)=="post_test"] <- "test"

pre20 <- subset(pre_summary, amt == 20)
post20 <- subset(post_summary, amt == 20)

shuffled_pre20 <- subset(shuffled_pre_summary, amt == 20)
shuffled_post20 <- subset(shuffled_post_summary, amt == 20)

# # Merge pre20 and post20
pre20$merge1 <- paste(pre20$sentid, pre20$csize, pre20$clist, pre20$nhid, pre20$adaptlist, pre20$test, sep = '')

post20$merge1 <- paste(post20$sentid, post20$csize, post20$clist, post20$nhid, post20$adaptlist, post20$test, sep = '')

all20 <- merge(post20, pre20, by='merge1')

shuffled_pre20$merge1 <- paste(shuffled_pre20$sentid, shuffled_pre20$csize, shuffled_pre20$clist, shuffled_pre20$nhid, shuffled_pre20$adaptlist, shuffled_pre20$test, sep = '')

shuffled_post20$merge1 <- paste(shuffled_post20$sentid, shuffled_post20$csize, shuffled_post20$clist, shuffled_post20$nhid, shuffled_post20$adaptlist, shuffled_post20$test, sep = '')

shuffled_all20 <- merge(shuffled_post20, shuffled_pre20, by='merge1')

colnames(all20)[2:8] <- c('sentid', 'csize', 'clist', 'nhid', 'adaptlist', 'test', 'amt')
drop <- c('sentid.y', 'csize.y', 'clist.y', 'nhid.y', 'adaptlist.y', 'test.y', 'amt.y')
all20 <- all20[,!(names(all20) %in% drop)]

colnames(shuffled_all20)[2:8] <- c('sentid', 'csize', 'clist', 'nhid', 'adaptlist', 'test', 'amt')
drop <- c('sentid.y', 'csize.y', 'clist.y', 'nhid.y', 'adaptlist.y', 'test.y', 'amt.y')
shuffled_all20 <- shuffled_all20[,!(names(shuffled_all20) %in% drop)]

#add acceptablity
leftbranch <- c(1, 2, 5, 6, 14, 19, 22, 23)
coordstruct <- c(1, 3, 4, 6, 7, 11, 14, 22, 23)

all20$acceptability <- 0
for (row in 1:nrow(all20)){
  construction <- all20[row, "test"]
  id <- all20[row, "sentid"]
  if(construction=='adjunct') {all20[row, "acceptability"] = 1}
  if(construction=='wh') {all20[row, "acceptability"] = 1}
  if(construction=='subject') {all20[row, "acceptability"] = 1}
  if(construction=='object') {all20[row, "acceptability"] = 5}
  if(construction=='leftbranch') {all20[row, "acceptability"] = 1}
  if(construction=='leftbranch' && id %in% leftbranch) {all20[row, "acceptability"] = 5}
  if(construction=='coordstruct') {all20[row, "acceptability"] = 1}
  if(construction=='coordstruct' && id %in% coordstruct) {all20[row, "acceptability"] = 5}
  if(construction=='complexnp') {all20[row, "acceptability"] = 1}
  if(construction=='bridge') {all20[row, "acceptability"] = 5}
}

shuffled_all20$acceptability <- 0
for (row in 1:nrow(shuffled_all20)){
  construction <- shuffled_all20[row, "test"]
  id <- shuffled_all20[row, "sentid"]
  if(construction=='adjunct') {shuffled_all20[row, "acceptability"] = 1}
  if(construction=='wh') {shuffled_all20[row, "acceptability"] = 1}
  if(construction=='subject') {shuffled_all20[row, "acceptability"] = 1}
  if(construction=='object') {shuffled_all20[row, "acceptability"] = 5}
  if(construction=='leftbranch') {shuffled_all20[row, "acceptability"] = 1}
  if(construction=='leftbranch' && id %in% leftbranch) {shuffled_all20[row, "acceptability"] = 5}
  if(construction=='coordstruct') {shuffled_all20[row, "acceptability"] = 1}
  if(construction=='coordstruct' && id %in% coordstruct) {shuffled_all20[row, "acceptability"] = 5}
  if(construction=='complexnp') {shuffled_all20[row, "acceptability"] = 1}
  if(construction=='bridge') {shuffled_all20[row, "acceptability"] = 5}
}

# # Add relevant diff columns

strucs = c('adjunct', 'wh', 'subject', 'object', 'leftbranch', 'coordstruct', 'complexnp', 'bridge')

# Reordering levels for correct RSM plotting
all20$test <- factor(all20$test, levels = strucs)
all20$adapt = factor(all20$adapt, levels = strucs)

shuffled_all20$test <- factor(shuffled_all20$test, levels = strucs)
shuffled_all20$adapt = factor(shuffled_all20$adapt, levels = strucs)


#ttestall <- merge(all20, shuffled_all20, by='merge1')

```



### Add in useful variables
```{r, cache=TRUE}

#all20$diagonal <- ifelse(all20$adapt == all20$test, 'diagonal', 'not-diagonal')

# rcs <- c('orc','orrc','prc','prrc','src')
# all20$testRC <- ifelse(all20$test %in% rcs, 'Test on RCs', 'Test on coordination')
# all20$testRC <- factor(all20$testRC, levels = c('Test on RCs', 'Test on coordination'))
# all20$adaptRC <- ifelse(all20$adapt %in% rcs, 'Adapted to RCs', 'Adapted to coordination')
# all20$adaptRC <- factor(all20$adaptRC, levels = c('Adapted to RCs', 'Adapted to coordination'))

# linked <- c('linkedwh')
# 
# all20$testLinked <- ifelse(all20$test %in% linked, 'Test on D-Linked', 'Test on non D-Linked')
# all20$testLinked <- factor(all20$testLinked, levels = c('Test on D-Linked', 'Test on non D-Linked'))
# 
# all20$adaptLinked <- ifelse(all20$adapt %in% linked, 'Adapted on D-Linked', 'Adapted on non D-Linked')
# all20$adaptLinked <- factor(all20$adaptLinked, levels = c('Adapted on D-Linked', 'Adapted on non D-Linked'))
# 
# shuffled_all20$testLinked <- ifelse(shuffled_all20$test %in% linked, 'Test on D-Linked', 'Test on non D-Linked')
# shuffled_all20$testLinked <- factor(shuffled_all20$testLinked, levels = c('Test on D-Linked', 'Test on non D-Linked'))
# 
# shuffled_all20$adaptLinked <- ifelse(shuffled_all20$adapt %in% linked, 'Adapted on D-Linked', 'Adapted on non D-Linked')
# shuffled_all20$adaptLinked <- factor(shuffled_all20$adaptLinked, levels = c('Adapted on D-Linked', 'Adapted on non D-Linked'))

valid <- c('object', 'bridge', 'coordstruct', 'leftbranch')

all20$testValid <- ifelse(all20$acceptability == 5, 'Test on Acceptable', 'Test on Unacceptable')
all20$testValid <- factor(all20$testValid, levels = c('Test on Acceptable', 'Test on Unacceptable'))

all20$adaptValid <- ifelse(all20$adapt %in% valid, 'Adapted on Acceptable', 'Adapted on Unacceptable')
all20$adaptValid <- factor(all20$adaptValid, levels = c('Adapted on Acceptable', 'Adapted on Unacceptable'))

shuffled_all20$testValid <- ifelse(shuffled_all20$acceptability == 5, 'Test on Acceptable', 'Test on Unacceptable')
shuffled_all20$testValid <- factor(shuffled_all20$testValid, levels = c('Test on Acceptable', 'Test on Unacceptable'))

shuffled_all20$adaptValid <- ifelse(shuffled_all20$adapt %in% valid, 'Adapted on Acceptable', 'Adapted on Unacceptable')
shuffled_all20$adaptValid <- factor(shuffled_all20$adaptValid, levels = c('Adapted on Acceptable', 'Adapted on Unacceptable'))

sub<- c('wh', 'subject', 'complexnp', 'leftbranch')

all20$testSub <- ifelse(all20$test %in% sub, 'Test on Subjacency', 'Test on not subjacency')
all20$testSub <- factor(all20$testSub, levels = c('Test on Subjacency', 'Test on not subjacency'))

all20$adaptSub <- ifelse(all20$adapt %in% sub, 'Adapted on Subjacency', 'Adapted on not subjacency')
all20$adaptSub <- factor(all20$adaptSub, levels = c('Adapted on Subjacency', 'Adapted on not subjacency'))

shuffled_all20$testSub <- ifelse(shuffled_all20$test %in% sub, 'Test on Subjacency', 'Test on not subjacency')
shuffled_all20$testSub <- factor(shuffled_all20$testSub, levels = c('Test on Subjacency', 'Test on not subjacency'))

shuffled_all20$adaptSub <- ifelse(shuffled_all20$adapt %in% sub, 'Adapted on Subjacency', 'Adapted on not subjacency')
shuffled_all20$adaptSub <- factor(shuffled_all20$adaptSub, levels = c('Adapted on Subjacency', 'Adapted on not subjacency'))

Lmarking <- c('adjunct', 'subject', 'object')
Lmarking_neg <- c('adjunct', 'subject')
Lmarking_pos <- c('object')

all20$testLmarking <- ifelse(all20$test %in% Lmarking, 'Test on Lmark', 'Test on not Lmark')
all20$testLmarking <- factor(all20$testLmarking, levels = c('Test on Lmark', 'Test on not Lmark'))

all20$adaptLmarking <- ifelse(all20$adapt %in% Lmarking, 'Adapted on Lmark', 'Adapted on not Lmark')
all20$adaptLmarking <- factor(all20$adaptLmarking, levels = c('Adapted on Lmark', 'Adapted on not Lmark'))

shuffled_all20$testLmarking <- ifelse(shuffled_all20$test %in% Lmarking, 'Test on Lmark', 'Test on not Lmark')
shuffled_all20$testLmarking <- factor(shuffled_all20$testLmarking, levels = c('Test on Lmark', 'Test on not Lmark'))

shuffled_all20$adaptLmarking <- ifelse(shuffled_all20$adapt %in% Lmarking, 'Adapted on Lmark', 'Adapted on not Lmark')
shuffled_all20$adaptLmarking <- factor(shuffled_all20$adaptLmarking, levels = c('Adapted on Lmark', 'Adapted on not Lmark'))

empty <- c('subject', 'adjunct', 'object', 'coordstruct')
empty_neg <- c('subject', 'adjunct')
empty_pos <- c('object', 'coordstruct')

all20$testEmpty <- ifelse(all20$test %in% empty, 'Test on ECP', 'Test on not ECP')
all20$testEmpty <- factor(all20$testEmpty, levels = c('Test on ECP', 'Test on not ECP'))

all20$adaptEmpty <- ifelse(all20$adapt %in% empty, 'Adapted on ECP', 'Adapted on not ECP')
all20$adaptEmpty <- factor(all20$adaptEmpty, levels = c('Adapted on ECP', 'Adapted on not ECP'))

shuffled_all20$testEmpty <- ifelse(shuffled_all20$test %in% empty, 'Test on ECP', 'Test on not ECP')
shuffled_all20$testEmpty <- factor(shuffled_all20$testEmpty, levels = c('Test on ECP', 'Test on not ECP'))

shuffled_all20$adaptEmpty <- ifelse(shuffled_all20$adapt %in% empty, 'Adapted on ECP', 'Adapted on not ECP')
shuffled_all20$adaptEmpty <- factor(shuffled_all20$adaptEmpty, levels = c('Adapted on ECP', 'Adapted on not ECP'))

Dlinking <- c('wh', 'leftbranch', 'object')

all20$testDlinking <- ifelse(all20$test %in% Dlinking, 'Test on Dlink', 'Test on not Dlink')
all20$testDlinking <- factor(all20$testDlinking, levels = c('Test on Dlink', 'Test on not Dlink'))

all20$adaptDlinking <- ifelse(all20$adapt %in% Dlinking, 'Adapted on Dlink', 'Adapted on not Dlink')
all20$adaptDlinking <- factor(all20$adaptDlinking, levels = c('Adapted on Dlink', 'Adapted on not Dlink'))

shuffled_all20$testDlinking <- ifelse(shuffled_all20$test %in% Dlinking, 'Test on Dlink', 'Test on not Dlink')
shuffled_all20$testDlinking <- factor(shuffled_all20$testDlinking, levels = c('Test on Dlink', 'Test on not Dlink'))

shuffled_all20$adaptDlinking <- ifelse(shuffled_all20$adapt %in% Dlinking, 'Adapted on Dlink', 'Adapted on not Dlink')
shuffled_all20$adaptDlinking <- factor(shuffled_all20$adaptDlinking, levels = c('Adapted on Dlink', 'Adapted on not Dlink'))


```

## Calculating the adaptation effect

```{r, cache = TRUE}

# Adapt(Y | X)
all20$diff <- all20$pre_surp - all20$post_surp   

shuffled_all20$diff <- shuffled_all20$pre_surp - shuffled_all20$post_surp   

## Plotting the relationship between diff and pre_surp

p <- ggplot(all20, aes(x = diff, y = pre_surp)) + geom_point() + geom_smooth(method='lm') + labs(x = 'A(Y|X)', y = 'Surp(Y)')
p

p <- ggplot(shuffled_all20, aes(x = diff, y = pre_surp)) + geom_point() + geom_smooth(method='lm') + labs(x = 'A(Y|X)', y = 'Surp(Y)')
p

#ggsave('presurp_diff.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 7, height = 3.5, units = "in")

# Fitting model for Adapt(Y | X) ~ Surp(Y)
model <- lm(diff ~ scale(pre_surp, scale = FALSE), data = all20)
summary(model)

shuffled_model <- lm(diff ~ scale(pre_surp, scale = FALSE), data = shuffled_all20)
summary(shuffled_model)

# AE(Y | X)
all20$corrected_diff <- residuals(model) + coef(model)["(Intercept)"]

shuffled_all20$corrected_diff <- residuals(shuffled_model) + coef(shuffled_model)["(Intercept)"]

ttestall <- merge(all20, shuffled_all20, by=c('merge1', 'adapt', 'test'))

```
```{r}
#reproducing python analysis 

all20$Dlink_effect <- ifelse(all20$adapt %in% Dlinking & all20$test %in% Dlinking & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$Dlink_comp_effect <- ifelse(!(all20$adapt %in% Dlinking) & all20$test %in% Dlinking & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$Subjacency_effect <- ifelse(all20$adapt %in% sub & all20$test %in% sub & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$Subjacency_comp_effect <- ifelse(!(all20$adapt %in% sub) & all20$test %in% sub & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$Lmark_effect <- ifelse(((all20$adapt %in% Lmarking_neg & all20$test %in% Lmarking_neg)|(all20$adapt %in% Lmarking_pos & all20$test %in% Lmarking_pos)) & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$Lmark_comp_effect <- ifelse(!(all20$adapt %in% Lmarking) & all20$test %in% Lmarking & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$Lmark_blue_effect <- ifelse(((all20$adapt %in% Lmarking_neg & all20$test %in% Lmarking_pos)|(all20$adapt %in% Lmarking_pos & all20$test %in% Lmarking_neg)) & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$ECP_effect <- ifelse(((all20$adapt %in% empty_neg & all20$test %in% empty_neg)|(all20$adapt %in% empty_pos & all20$test %in% empty_pos)) & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$ECP_comp_effect <- ifelse(!(all20$adapt %in% empty) & all20$test %in% empty & all20$test != all20$adapt, all20$corrected_diff, NaN)

all20$ECP_blue_effect <- ifelse(((all20$adapt %in% empty_neg & all20$test %in% empty_pos)|(all20$adapt %in% empty_pos & all20$test %in% empty_neg)) & all20$test != all20$adapt, all20$corrected_diff, NaN)
  
  
```
```{r}
#t-tests

t.test(all20$Dlink_effect, all20$Dlink_comp_effect)
t.test(all20$Subjacency_effect, all20$Subjacency_comp_effect)
t.test(all20$Lmark_effect, all20$Lmark_comp_effect)
t.test(all20$ECP_effect, all20$ECP_comp_effect)
t.test(all20$ECP_effect, all20$ECP_blue_effect)
t.test(all20$Lmark_effect, all20$Lmark_blue_effect)

#sum(!is.na(all20$Lmark_effect))

```


### Analysis 1: Validating AE as a similarity metric

#### Creating Figure 2a

```{r, cache=TRUE}

labs <- c(
'Adjunct extraction',
'Wh extraction',
'Subject extraction',
'Object extraction',
'Left branch extraction',
'Coordinate Structure extraction',
'Complex NP extraction',
'Bridge Verb extraction'
)

summ_analysis1 <- data.frame(NULL)
summ_analysis2 <- data.frame(NULL)

# for(struc in strucs){
#   curr <- subset(all20, adapt == struc)
#   curr$structure <- ifelse(curr$test == struc, 'same', 'different')
#   curr$structure <- factor(curr$structure, levels = c('same', 'different'))
#   curr_summ <- data_summary(curr, 'corrected_diff', quos(structure))
#   curr_summ$struc <- struc
#   summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# }
# summ_analysis1$struc <- factor(summ_analysis1$struc, levels = rev(c('adjunct', 'wh', 'subject', 'object','leftbranch', 'coordstruct','complexnp', 'bridge')))
# p <- ggplot(summ_analysis1, aes(struc,corrected_diff, fill = structure, group = structure)) + geom_bar(stat='identity', width=0.5, position=position_dodge(width=0.5))   + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.5))  + coord_flip() + scale_x_discrete(labels = rev(labs)) + labs(x='', y='Adaptation Effect \n (bits of surprisal)', fill = 'Test structure') +  scale_fill_manual(values = c('plum1', 'royalblue3'), labels = c('Same as adaptation', 'Different from adaptation'))
# p

# curr <- subset(all20, adaptValid == 'Adapted on Acceptable' & testValid ==  'Test on Acceptable')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptValid))
# curr_summ$struc <- 'same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(all20, adaptValid == 'Adapted on Unacceptable' & testValid == 'Test on Unacceptable')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptValid))
# curr_summ$struc <- 'same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(all20, adaptValid == 'Adapted on Acceptable' & testValid ==  'Test on Unacceptable')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptValid))
# curr_summ$struc <- 'different'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(all20, adaptValid == 'Adapted on Unacceptable' & testValid == 'Test on Acceptable')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptValid))
# curr_summ$struc <- 'different'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(shuffled_all20, adaptValid == 'Adapted on Acceptable' & testValid ==  'Test on Acceptable')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptValid))
# curr_summ$struc <- 'shuffled same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(shuffled_all20, adaptValid == 'Adapted on Unacceptable' & testValid == 'Test on Unacceptable')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptValid))
# curr_summ$struc <- 'shuffled same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)


# curr <- subset(all20, adaptLinked == 'Adapted on D-Linked' & testLinked ==  'Test on D-Linked')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLinked))
# curr_summ$struc <- 'same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
#
# curr <- subset(all20, adaptLinked == 'Adapted on non D-Linked' & testLinked == 'Test on non D-Linked')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLinked))
# curr_summ$struc <- 'same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
#
# curr <- subset(all20, adaptLinked == 'Adapted on D-Linked' & testLinked ==  'Test on non D-Linked')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLinked))
# curr_summ$struc <- 'different'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
#
# curr <- subset(all20, adaptLinked == 'Adapted on non D-Linked' & testLinked == 'Test on D-Linked')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLinked))
# curr_summ$struc <- 'different'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
#
# curr <- subset(shuffled_all20, adaptLinked == 'Adapted on D-Linked' & testLinked ==  'Test on D-Linked')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLinked))
# curr_summ$struc <- 'shuffled same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
#
# curr <- subset(shuffled_all20, adaptLinked == 'Adapted on non D-Linked' & testLinked == 'Test on non D-Linked')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLinked))
# curr_summ$struc <- 'shuffled same'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)

# #generating graph for L-marking constraint

curr <- subset(all20, test == adapt & adaptLmarking == 'Adapted on Lmark')
curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLmarking))
curr_summ$struc <- 'same construction'
summ_analysis1 <- rbind(summ_analysis1, curr_summ)

curr <- subset(all20, test != adapt & adaptLmarking == 'Adapted on Lmark' & (adapt %in% Lmarking_neg & test %in% Lmarking_neg) | (adapt %in% Lmarking_pos & test %in% Lmarking_pos))
curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLmarking))
curr_summ$struc <- 'same adherence within constraint'
summ_analysis1 <- rbind(summ_analysis1, curr_summ)

curr <- subset(all20, test != adapt & adaptLmarking == 'Adapted on Lmark' & (adapt %in% Lmarking_neg & test %in% Lmarking_pos) | (adapt %in% Lmarking_pos & test %in% Lmarking_neg))
curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLmarking))
curr_summ$struc <- 'opposite adherence within constraint'
summ_analysis1 <- rbind(summ_analysis1, curr_summ)

curr <- subset(all20, test != adapt & !(adapt %in% Lmarking) & test %in% Lmarking)
curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLmarking))
curr_summ$struc <- 'outside constraint'
summ_analysis1 <- rbind(summ_analysis1, curr_summ)

# curr <- subset(shuffled_all20, test != adapt & adaptLmarking == 'Adapted on Lmark' & (adapt %in% Lmarking_neg & test %in% Lmarking_neg) | (adapt %in% Lmarking_pos & test %in% Lmarking_pos))
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLmarking))
# curr_summ$struc <- 'shuffled same adherence within constraint'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(shuffled_all20, test != adapt & adaptLmarking == 'Adapted on Lmark' & (adapt %in% Lmarking_neg & test %in% Lmarking_pos) | (adapt %in% Lmarking_pos & test %in% Lmarking_neg))
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptLmarking))
# curr_summ$struc <- 'shuffled opposite adherence within constraint'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)

#generating graph for Empty Category Principle constraint
curr2 <- subset(all20, test == adapt & adaptEmpty == 'Adapted on ECP')
curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptEmpty))
curr_summ2$struc <- 'same construction'
summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

curr2 <- subset(all20, test != adapt & adaptEmpty == 'Adapted on ECP' & ((adapt %in% empty_neg & test %in% empty_neg) | (adapt %in% empty_pos & test %in% empty_pos)))
curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptEmpty))
curr_summ2$struc <- 'same adherence within constraint'
summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

curr2 <- subset(all20, test != adapt & adaptEmpty == 'Adapted on ECP' & ((adapt %in% empty_neg & test %in% empty_pos) | (adapt %in% empty_pos & test %in% empty_neg)))
curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptEmpty))
curr_summ2$struc <- 'opposite adherence within constraint'
summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

curr2 <- subset(all20, test != adapt & !(adapt %in% empty) & test %in% empty)
curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptEmpty))
curr_summ2$struc <- 'outside constraint'
summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

# curr2 <- subset(shuffled_all20, test != adapt & adaptEmpty == 'Adapted on ECP' & ((adapt %in% empty_neg & test %in% empty_neg) | (adapt %in% empty_pos & test %in% empty_pos)))
# curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptEmpty))
# curr_summ2$struc <- 'shuffled same adherence within constraint'
# summ_analysis2 <- rbind(summ_analysis2, curr_summ2)
# 
# curr2 <- subset(shuffled_all20, test != adapt & adaptEmpty == 'Adapted on ECP' & ((adapt %in% empty_neg & test %in% empty_pos) | (adapt %in% empty_pos & test %in% empty_neg)))
# curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptEmpty))
# curr_summ2$struc <- 'shuffled opposite adherence within constraint'
# summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

# #generating graph for D-Linking constraint
# curr <- subset(all20, test == adapt & adaptDlinking == 'Adapted on Dlink')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptDlinking))
# curr_summ$struc <- 'same construction'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(all20, test != adapt & adaptDlinking == 'Adapted on Dlink')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptDlinking))
# curr_summ$struc <- 'different construction within constraint'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)
# 
# curr <- subset(all20, test != adapt & !(adapt %in% Dlinking) & test %in% Dlinking)
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptDlinking))
# curr_summ$struc <- 'outside constraint'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)

# curr <- subset(shuffled_all20, test != adapt & adaptDlinking == 'Adapted on Dlink')
# curr_summ <- data_summary(curr, 'corrected_diff', quos(adaptDlinking))
# curr_summ$struc <- 'shuffled different construction within constraint'
# summ_analysis1 <- rbind(summ_analysis1, curr_summ)

# #generating graph for subjacency constraint
# curr2 <- subset(all20, test == adapt & adaptSub == 'Adapted on Subjacency')
# curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptSub))
# curr_summ2$struc <- 'same construction'
# summ_analysis2 <- rbind(summ_analysis2, curr_summ2)
# 
# curr2 <- subset(all20, test != adapt & adaptSub == 'Adapted on Subjacency')
# curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptSub))
# curr_summ2$struc <- 'different construction within constraint'
# summ_analysis2 <- rbind(summ_analysis2, curr_summ2)
# 
# curr2 <- subset(all20, test != adapt & !(adapt %in% sub) & test %in% sub)
# curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptSub))
# curr_summ2$struc <- 'outside constraint'
# summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

# curr2 <- subset(shuffled_all20, test != adapt & adaptSub == 'Adapted on Subjacency')
# curr_summ2 <- data_summary(curr2, 'corrected_diff', quos(adaptSub))
# curr_summ2$struc <- 'shuffled different construction within constraint'
# summ_analysis2 <- rbind(summ_analysis2, curr_summ2)

# #summ_analysis1$struc <- factor(summ_analysis1$struc, levels = rev(c('adjunct', 'wh', 'subject', 'object','leftbranch', 'coordstruct','complexnp', 'bridge')))

# q <- ggplot(summ_analysis2, aes(adaptSub,corrected_diff, fill = struc, group = struc)) + geom_bar(stat='identity', width=0.5, position=position_dodge(width=0.5)) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.5))  + coord_flip() + labs(x='', y='Subjacency Adaptation Effect \n (bits of surprisal)', fill = 'Test structure') +  scale_fill_manual(values = c('purple','red', 'plum1'))
# q <- q+scale_x_discrete(breaks = seq(0, 2, by = 1))
# 
# p <- ggplot(summ_analysis1, aes(adaptDlinking,corrected_diff, fill = struc, group = struc)) + geom_bar(stat='identity', width=0.5, position=position_dodge(width=0.5)) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.5))  + coord_flip() + labs(x='', y='D-Linking Adaptation Effect \n (bits of surprisal)', fill = 'Test structure') +  scale_fill_manual(values = c('purple', 'red', 'plum1'))
# p <- p+scale_x_discrete(breaks = seq(0, 2, by = 1))
# # #
q <- ggplot(summ_analysis2, aes(adaptEmpty,corrected_diff, fill = struc, group = struc)) + geom_bar(stat='identity', width=0.5, position=position_dodge(width=0.5))  + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.5))  + coord_flip() + labs(x='', y='Empty Category Principle \n Adaptation Effect \n (bits of surprisal)', fill = 'Test structure') +  scale_fill_manual(values = c('royalblue3', 'red', 'purple', 'plum1'))
q <- q+scale_x_discrete(breaks = seq(0, 2, by = 1))

p <- ggplot(summ_analysis1, aes(adaptLmarking,corrected_diff, fill = struc, group = struc)) + geom_bar(stat='identity', width=0.5, position=position_dodge(width=0.5)) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.5))  + coord_flip() + labs(x='', y='L-marking Adaptation Effect \n (bits of surprisal)', fill = 'Test structure') +  scale_fill_manual(values = c('royalblue3', 'red', 'purple', 'plum1'))
p <- p+scale_x_discrete(breaks = seq(0, 2, by = 1))
# # 
library(grid)
grid.newpage()
grid.draw(rbind(ggplotGrob(p), ggplotGrob(q), size = "last"))

# p <- ggplot(summ_analysis1, aes(x=adaptValid,y=corrected_diff, fill = struc, group = struc)) + geom_bar(stat='identity', width=0.5, position=position_dodge(width=0.5)) + geom_text(aes(label = corrected_diff), position=position_dodge(width=0.9)) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.5))  + coord_flip() + labs(x='', y='Filler-Gap Adaptation Effect \n (bits of surprisal)', fill = 'Test structure') +  scale_fill_manual(values = c('royalblue3', 'plum1', 'red'))
# 
# p


#ggsave('analysis1.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 7, height = 3.5, units = "in")

```


#### Running statistical analyses

```{r, cache=TRUE}

all_analysis1 <- data.frame(NULL)

strucs = c('adjunct', 'wh', 'subject', 'object','leftbranch', 'coordstruct','complexnp', 'bridge')

for(struc in strucs){
  curr <- subset(all20, adapt == struc)
  curr$structure <- ifelse(curr$test == struc, 'same', 'different')
  curr$structure <- factor(curr$structure, levels = c('same', 'different'))

  contrasts(curr$structure) <- "contr.sum"

  #curr_model <- lmer(corrected_diff ~ structure + (1 | adaptlist) + (1 | clist), curr)
  curr_model <- lm(corrected_diff ~ structure, curr)

  curr_ests <- c(coef(summary(curr_model))[, "Estimate"], coef(summary(curr_model))[, "Std. Error"], coef(summary(curr_model))[, "t value"], coef(summary(curr_model))[, "Pr(>|t|)"])

  all_analysis1 <- rbind(all_analysis1, curr_ests)
}

colnames(all_analysis1) <- c('intercept.est', 'structure.est', 'intercept.se','structure.se', 'intercept.tval', 'structure.tval', 'intercept.pval', 'structure.pval')

all_analysis1$struc <- strucs

print(paste(all_analysis1$struc, all_analysis1$structure, round(all_analysis1$structure.est, 3), round(all_analysis1$structure.se, 3), all_analysis1$structure.pval))
#stop here
``` 


### Analysis 2: Similarity between sentences with different types of VP coordination 

#### Creating Figure 2b upper panel
```{r, cache = TRUE}

no_diagonal <- subset(all20, diagonal != 'diagonal')

no_diagonal$sameRC <- factor(ifelse(no_diagonal$adapt %in% rcs & no_diagonal$test %in% rcs, 'same', 
                       ifelse(!(no_diagonal$adapt %in% rcs | no_diagonal$test %in% rcs), 'same', 'different')), levels = c('same', 'different'))

no_diagonal$adaptRC <- factor(ifelse(no_diagonal$adapt %in% rcs, 'Adapt on RC', 'Adapt on coordination'), levels = c('Adapt on RC', 'Adapt on coordination'))

summ_analysis2_no.diagonal <- data_summary(no_diagonal, 'corrected_diff', .(sameRC, adaptRC, testRC))

ggplot(subset(summ_analysis2_no.diagonal, adaptRC == 'Adapt on coordination'), aes(testRC, corrected_diff, group = testRC, fill = sameRC)) +  geom_bar(stat='identity', width=0.5) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.9)) + theme(legend.position = "none") + labs(x = '', y = 'Adaptation effect \n (bits of surprisal)') + coord_flip() + scale_fill_manual(values = c('royalblue3', 'plum1'))

```

#### Statistical analyses

```{r, cache=TRUE}

contrasts(no_diagonal$testRC) <- "contr.sum"
contrasts(no_diagonal$testRC)

analysis2 <- lmer(corrected_diff ~ testRC +  (1 | adaptlist) + (1 | clist), subset(no_diagonal, !(adapt %in% rcs)))
summary(analysis2)

```

### Analysis 3: Similarity between sentences with different types of RCs 

#### Creating Figure 2b lower panel

```{r, cache=TRUE}

ggplot(subset(summ_analysis2_no.diagonal, adaptRC == 'Adapt on RC'), aes(testRC, corrected_diff, group = testRC, fill = sameRC)) +  geom_bar(stat='identity', width=0.5) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.9)) + theme(legend.position = "none") + labs(x = '', y = 'Adaptation effect \n (bits of surprisal)') + coord_flip() + scale_fill_manual(values = c('royalblue3', 'plum1'))

```


#### Statistical analyses

```{r, cache=TRUE}

contrasts(no_diagonal$testRC)

analysis3 <- lmer(corrected_diff ~ testRC +  (1 | adaptlist) + (1 | clist), subset(no_diagonal, adapt %in% rcs))
summary(analysis3)

```

```{r, include = FALSE}

## Saving Fig 2b

summ_analysis2_no.diagonal$adaptRC <- factor(summ_analysis2_no.diagonal$adaptRC, levels = c('Adapt on coordination', 'Adapt on RC'))

p <- ggplot(summ_analysis2_no.diagonal, aes(testRC, corrected_diff, group = testRC, fill = sameRC)) +  geom_bar(stat='identity', width=0.5) + geom_errorbar(aes(ymin=corrected_diff - ci, ymax=corrected_diff + ci),  width = 0.2, position = position_dodge(0.9)) + theme(legend.position = "none") + labs(x = '', y = 'Adaptation effect \n (bits of surprisal)') + coord_flip() + facet_wrap(~adaptRC, nrow=2) + scale_fill_manual(values = c('royalblue3', 'plum1'))

p

#ggsave('analysis2.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 7, height = 3.5, units = "in")

```

### Analysis 4: Similarity between sentences belonging to different sub-classes of RCs

#### Creating Figure 3
```{r, cache=TRUE}

passive_object <- subset(all20, adapt %in% c('prc', 'prrc', 'orc', 'orrc') & test %in% c('prc', 'prrc', 'orc', 'orrc'))

passive_object$adapt.passive <- ifelse(passive_object$adapt %in% c('prc', 'prrc'), 'passive', 'object')
passive_object$test.passive <- ifelse(passive_object$test %in% c('prc', 'prrc'), 'passive', 'object')
passive_object$adapt.reduced <- ifelse(passive_object$adapt %in% c('prc', 'orc'), 'Unreduced RC', 'Reduced RC')
passive_object$test.reduced <- ifelse(passive_object$test %in% c('prc', 'orc'), 'Unreduced RC', 'Reduced RC')

passive_object$isPassive <- ifelse(passive_object$adapt.passive == passive_object$test.passive, 'Match', 'Mismatch')

passive_object$isReduced <- ifelse(passive_object$adapt.reduced == passive_object$test.reduced, 'Match', 'Mismatch')

passive_object_summ <- data_summary(passive_object, 'corrected_diff', .(isPassive, isReduced))
passive_object_summ$isPassive <- factor(passive_object_summ$isPassive, levels = c('Match', 'Mismatch'))
passive_object_summ$isReduced <- factor(passive_object_summ$isReduced, levels = rev(c('Match', 'Mismatch')))

p <- ggplot(passive_object_summ, mapping = aes(x=isPassive, y=isReduced)) + geom_tile(aes(fill=corrected_diff)) +scale_fill_gradient(low = 'white',  high = 'steelblue') + labs(fill = 'Adaptation\nEffect', x='Passivity', y='Reduction') + geom_text(aes(label=round(corrected_diff, 2)))

p

#ggsave('analysis4.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 5, height = 3.5, units = "in")
```


#### Statistical analyses
```{r, cache=TRUE}

df <- subset(all20, test == 'src' & adapt %in% c('prrc', 'orrc', 'prc', 'orc'))

df_summ <- data_summary(df, 'corrected_diff', .(adapt))

graph <- ggplot(df_summ, mapping=aes(x=adapt, y=corrected_diff)) + geom_bar(stat='identity') + geom_errorbar(aes(ymin=corrected_diff-sd, ymax=corrected_diff+sd))

graph

passive_object$adapttest <- paste(passive_object$adapt, passive_object$test, sep = '')

passive_object$testtype <-  ifelse(passive_object$adapttest %in% c('prrcprrc', 'orrcorrc','prcprc', 'orcorc'), 'both match', 
                                   ifelse(passive_object$adapttest %in% c('prrcorrc', 'orrcprrc','prcorc', 'orcprc'), 'reduced match', 
                                          ifelse(passive_object$adapttest %in% c('prrcprc', 'orrcorc','prcprrc', 'orcorrc'), 'passive match', 'no match' )))

passive_object$testtype <- factor(passive_object$testtype, levels = c('passive match', 'reduced match', 'both match', 'no match'))

contrasts(passive_object$testtype)

analysis4_combined <- lmer(corrected_diff ~ testtype + (1 | adaptlist) + (1 | clist), passive_object)

summary(analysis4_combined)
```


### Analysis 5: Effect of model size and training corpus size on the similarity between sentences

#### Loading in models with no training

```{r, cache=TRUE}

## Load data
zero_pre_summary <- readRDS('../data/dataframes/pre_summary_zero_nounk.Rda')
zero_post_summary <- readRDS('../data/dataframes/post_summary_zero_nounk.Rda')

zero_pre_summary <- subset(zero_pre_summary, adaptlist > 5)
colnames(zero_pre_summary)[colnames(zero_pre_summary)=="pre_test"] <- "test"

zero_post_summary <- subset(zero_post_summary, adaptlist > 5)
colnames(zero_post_summary)[colnames(zero_post_summary)=="post_test"] <- "test"

zero_pre_summary$merge1 <- paste(zero_pre_summary$sentid, zero_pre_summary$nhid, zero_pre_summary$adaptlist, zero_pre_summary$test, sep = '')

zero_post_summary$merge1 <- paste(zero_post_summary$sentid, zero_post_summary$nhid, zero_post_summary$adaptlist, zero_post_summary$test, sep = '')

all_zero <- merge(zero_pre_summary,zero_post_summary, by='merge1')

colnames(all_zero)[2:5] <- c('sentid',  'nhid', 'adaptlist', 'test')
drop <- c('sentid.y', 'csize.y', 'clist.y', 'nhid.y', 'adaptlist.y', 'test.y', 'amt.y')
all_zero <- all_zero[,!(names(all_zero) %in% drop)]

## Add variables
all_zero$testRC <- ifelse(all_zero$test %in% rcs, 'Test on RCs', 'Test on coordination')
all_zero$testRC <- factor(all_zero$testRC, levels = c('Test on RCs', 'Test on coordination'))

all_zero$adaptRC <- ifelse(all_zero$adapt %in% rcs, 'Adapted to RCs', 'Adapted to coordination')
all_zero$adaptRC <- factor(all_zero$adaptRC, levels = c('Adapted to RCs', 'Adapted to coordination'))

all_zero$diagonal <- ifelse(all_zero$adapt == all_zero$test, 'diagonal', 'not-diagonal')

## Calculate adaptation effect
all_zero$diff <- all_zero$pre_surp - all_zero$post_surp

model_zero <- lm(diff ~ scale(pre_surp, scale = FALSE), data = all_zero)
summary(model_zero)

# AE(Y | X)
all_zero$corrected_diff <- residuals(model_zero) + coef(model_zero)["(Intercept)"]

```


#### Creating Figure 5a

```{r, cache=TRUE}

# RC vs. not RC
summ_analysis5.1 <- subset(all20, adapt %in% rcs & diagonal != 'diagonal')  %>% group_by(testRC, nhid, csize, clist, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff") %>%  mutate(prop = `Test on RCs` / `Test on coordination`) %>%  group_by(nhid, csize) %>% summarise(prop = mean(prop))

summ_analysis5.1_zero <- subset(all_zero, adapt %in% rcs & diagonal != 'diagonal')  %>% group_by(testRC, nhid,adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff") %>%  mutate(prop = `Test on RCs` / `Test on coordination`) %>%  group_by(nhid) %>% summarise(prop = mean(prop))

summ_analysis5.1_zero$csize <- 0

summ_analysis5.1 <- bind_rows(summ_analysis5.1, summ_analysis5.1_zero)
summ_analysis5.1$adapttype <- 'All RCs'


# Match vs. mismatch in reduction
all20$adaptReduced <- ifelse(all20$adapt %in% c('orrc', 'prrc'), 'reduced', 'unreduced')
all20$testReduced <- ifelse(all20$test %in% c('orrc', 'prrc'), 'reduced', 'unreduced')

all20$reduced <- ifelse(all20$adaptReduced == all20$testReduced, 'match', 'mismatch')

all_zero$adaptReduced <- ifelse(all_zero$adapt %in% c('orrc', 'prrc'), 'reduced', 'unreduced')
all_zero$testReduced <- ifelse(all_zero$test %in% c('orrc', 'prrc'), 'reduced', 'unreduced')

all_zero$reduced <- ifelse(all_zero$adaptReduced == all_zero$testReduced, 'match', 'mismatch')

summ_analysis5.2 <- subset(all20, adapt %in% rcs) %>% group_by(reduced, nhid, csize, clist, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="reduced",value="corrected_diff") %>%  mutate(prop = `match` / `mismatch`) %>%  group_by(nhid, csize) %>% summarise(prop = mean(prop))

summ_analysis5.2_zero <- subset(all_zero, adapt %in% rcs) %>% group_by(reduced, nhid, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="reduced",value="corrected_diff") %>%  mutate(prop = `match` / `mismatch`) %>%  group_by(nhid) %>% summarise(prop = mean(prop))

summ_analysis5.2_zero$csize <- 0
summ_analysis5.2 <- bind_rows(summ_analysis5.2, summ_analysis5.2_zero)

summ_analysis5.2$adapttype <- 'Reduced RCs' 


# Specific RC

summ_analysis5.3 <- subset(all20, adapt %in% rcs)  %>% group_by(diagonal, nhid, csize, clist, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="diagonal",value="corrected_diff") %>%  mutate(prop = `diagonal` / `not-diagonal`) %>%  group_by(nhid, csize) %>% summarise(prop = mean(prop))

summ_analysis5.3_zero <- subset(all_zero, adapt %in% rcs)  %>% group_by(diagonal, nhid,adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="diagonal",value="corrected_diff") %>%  mutate(prop = `diagonal` / `not-diagonal`) %>%  group_by(nhid) %>% summarise(prop = mean(prop))

summ_analysis5.3_zero$csize <- 0

summ_analysis5.3 <- bind_rows(summ_analysis5.3, summ_analysis5.3_zero)
summ_analysis5.3$adapttype <- 'Specific RCs' 

summ_analysis5 <- rbind(summ_analysis5.1, summ_analysis5.2, summ_analysis5.3)
summ_analysis5$nhid <- factor(summ_analysis5$nhid)
summ_analysis5$csize <- factor(summ_analysis5$csize)

summ_analysis5$adapttype <- factor(summ_analysis5$adapttype, levels = c('Specific RCs', 'Reduced RCs', 'All RCs'))

purples = brewer.pal(n = 9, "Purples")[3:8]

p <- ggplot(subset(summ_analysis5) ,aes(x=csize, y = prop, fill = nhid)) + geom_bar(stat='identity',position='dodge') + scale_fill_manual(values = purples) + labs(x = 'Training corpus size \n (in millions of tokens)', y = 'D(Members, Non-members)', fill = 'Number of \nhidden units')  + facet_wrap(~adapttype) + geom_hline(yintercept = 1)

p

#ggsave('analysis5.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 7, height = 3.5, units = "in")

```


#### Statistical analyses

```{r, cache=TRUE}

# RC vs. not RC
analysis5.1 <- subset(all20, adapt %in% rcs & diagonal != 'diagonal')  %>% group_by(testRC, nhid, csize, clist, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff") %>%  mutate(prop = `Test on RCs` / `Test on coordination`) 

model5.1 <- lmer(prop~scale(nhid)*scale(csize) + (1|clist) + (1|adaptlist), data = analysis5.1)
summary(model5.1)

# Match vs. mismatch in reduction
analysis5.2 <- subset(all20, adapt %in% rcs) %>% group_by(reduced, nhid, csize, clist, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="reduced",value="corrected_diff") %>%  mutate(prop = `match` / `mismatch`) 

model5.2 <- lmer(prop~scale(nhid)*scale(csize) + (1|clist) + (1|adaptlist), data = analysis5.2)
summary(model5.2)

# Specific RC
analysis5.3 <- subset(all20, adapt %in% rcs)  %>% group_by(diagonal, nhid, csize, clist, adaptlist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="diagonal",value="corrected_diff") %>%  mutate(prop = `diagonal` / `not-diagonal`) 

model5.3 <- lmer(prop~scale(nhid)*scale(csize) + (1|clist) + (1|adaptlist), data = analysis5.3)
summary(model5.3)

```


### Analysis 6: Does D(RC, coordination) predict agreement prediction accuracy? 

#### Load in accuracy data
```{r, cache=TRUE}

acc_dat <- read.table('../data/all_acc.dat', header = TRUE)

acc_orc <- subset(acc_dat, syntax == 'obj_rel_across')
acc_orc$merge_var <- paste(acc_orc$corpus_size, acc_orc$d_model, acc_orc$corpus_var, sep = '')
#colnames(acc_orc)[colnames(acc_orc)=="accuracy"] <- "orc.acc"
  
acc_orrc <- subset(acc_dat, syntax == 'obj_rel_no_comp_across')
acc_orrc$merge_var <- paste(acc_orrc$corpus_size, acc_orrc$d_model, acc_orrc$corpus_var, sep = '')
#colnames(acc_orrc)[colnames(acc_orrc)=="accuracy"] <- "orrc.acc"

acc_src <- subset(acc_dat, syntax == 'subj_rel')
acc_src$merge_var <- paste(acc_src$corpus_size, acc_src$d_model, acc_src$corpus_var, sep = '')
#colnames(acc_src)[colnames(acc_src)=="accuracy"] <- "src.acc"


acc_scont <- subset(acc_dat, syntax == 'long_vp_coord')
acc_scont$merge_var <- paste(acc_scont$corpus_size, acc_scont$d_model, acc_scont$corpus_var, sep = '')

```


#### Creating Figure 5b

```{r, cache=TRUE}

## ORC


orc_models <- subset(all20, adapt == 'orc' & diagonal != 'diagonal') %>% group_by(testRC, nhid, csize, clist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff")  %>%  mutate(prop =  `Test on RCs`/`Test on coordination`)

orc_models$merge_var <- paste(orc_models$csize,'m',orc_models$nhid,orc_models$clist, sep = '')

orc_models <- merge(orc_models, acc_orc, by = 'merge_var')
orc_models$struc <- 'Unreduced Object RC'

## ORRC

orrc_models <- subset(all20, adapt == 'orrc' & diagonal != 'diagonal') %>% group_by(testRC, nhid, csize, clist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff")  %>%  mutate(prop =   `Test on RCs`/`Test on coordination`)

orrc_models$merge_var <- paste(orrc_models$csize,'m',orrc_models$nhid,orrc_models$clist, sep = '')

orrc_models <- merge(orrc_models, acc_orrc, by = 'merge_var')
orrc_models$struc <- 'Reduced Object RC'

alltypes <- rbind(orc_models, orrc_models)
alltypes$csize <- factor(alltypes$csize)
alltypes$nhid <- factor(alltypes$nhid)

p <- ggplot(subset(alltypes, struc != 'Subject RC'), aes(x = prop, y = accuracy, color = nhid,  shape = nhid)) + geom_point() + geom_smooth(method='lm', se=FALSE) + facet_wrap(~struc) + scale_color_manual(values = purples) +  labs(x = 'D(RC, non-RC)', y = 'Accuracy on agreement prediction task', color = 'Number of \nhidden units', shape = 'Number of \nhidden units', size = 'Number of hidden units') + ylim(0,1)
p

#ggsave('analysis6_2.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 7, height = 4.3, units = "in")

```



```{r, cache = TRUE}

model_orc <- lm(accuracy ~ prop + scale(nhid) + scale(csize), data = orc_models)
summary(model_orc)

model_orrc <- lm(accuracy ~ prop + scale(nhid) + scale(csize), data = orrc_models)
summary(model_orrc)

```


#### Other structures

```{r, cache=TRUE}

## SRC
src_models <- subset(all20, adapt == 'src' & diagonal != 'diagonal') %>% group_by(testRC, nhid, csize, clist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff")  %>%  mutate(prop =   `Test on RCs`/`Test on coordination`)

src_models$merge_var <- paste(src_models$csize,'m',src_models$nhid,src_models$clist, sep = '')

src_models <- merge(src_models, acc_src, by = 'merge_var')
src_models$struc <- 'Subject RC'

## SCONT

scont_models <- subset(all20, adapt == 'scont' & diagonal != 'diagonal') %>% group_by(testRC, nhid, csize, clist) %>% summarise(corrected_diff = mean(corrected_diff)) %>% spread(key="testRC",value="corrected_diff")  %>%  mutate(prop =  `Test on coordination`/`Test on RCs`)

scont_models$merge_var <- paste(scont_models$csize,'m',scont_models$nhid,scont_models$clist, sep = '')

scont_models <- merge(scont_models, acc_scont, by = 'merge_var')
scont_models$struc <- 'Long coordination'

alltypes <- rbind(src_models, scont_models)
alltypes$csize <- factor(alltypes$csize)
alltypes$nhid <- factor(alltypes$nhid)

p <- ggplot(alltypes, aes(x = prop, y = accuracy, color = nhid,  shape = nhid)) + geom_point() + geom_smooth(method='lm', se=FALSE) + facet_wrap(~struc) + scale_color_manual(values = purples) +  labs(x = 'D(RC, non-RC)', y = 'Accuracy on agreement prediction task', color = 'Number of \nhidden units', shape = 'Number of \nhidden units', size = 'Number of hidden units') + ylim(0,1)
p

#ggsave('src_scont_acc.pdf', plot = p, device = "pdf", path = '../conll-plots/', width = 7, height = 4.3, units = "in")

```

```{r, cache=TRUE}

model_src <- lm(accuracy ~ prop + scale(nhid) + scale(csize), data = src_models)
summary(model_src)

model_scont <- lm(accuracy ~ prop + scale(nhid) + scale(csize), data = scont_models)
summary(model_scont)

```
